---
title: "Assignment 9"
author: "Melanie Kazenel"
date: "March 22, 2017"
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

## 1. Modify your existing functions so that only the dataframe is the argument input

```{r}
##################################################
# FUNCTION: anova
# performs an analysis of variance
# inputs: a vector of categorical data (x) and a vector of continuous data (y)
# outputs: degrees of freedom, sums of squares, mean, F-value, and p-value
#------------------------------------------------- 
anova <- function(dataFrame=data.frame(as.factor(rep(c("High","Low"),each=5)),runif(10))) {
  x <- dataFrame[,1]
  y <- dataFrame[,2]
  newDataFrame <- data.frame(x,y)
  myMod <- aov(y~x, data = newDataFrame)
  myOut <- summary(myMod)
  return(myOut)
}

anova()

##################################################
# FUNCTION: anovaGraph
# performs an analysis of variance and creates a boxplot depicting the results
# inputs: a vector of categorical data (x) and a vector of continuous data (y)
# outputs: barplot 
#------------------------------------------------- 
anovaGraph <- function(dataFrame=data.frame(as.factor(rep(c("High","Low"),each=5)),runif(10))) {
  x <- dataFrame[,1]
  y <- dataFrame[,2]
  newDataFrame <- data.frame(x,y)
  myMod <- aov(y~x, data = newDataFrame)
  myOut <- boxplot(y~x,data=newDataFrame,col=c("deepskyblue4","deepskyblue3","deepskyblue2"))
  return(myOut)
}

anovaGraph()

```


## 2. Use your data from HW4 to input into your functions (where applicable) and produce output and plots
```{r}

# Read in the data from Homework 4
z <- read.table("Data.csv", header = TRUE, sep=",", stringsAsFactors=FALSE)

# Effect of warming treatment on Shannon diversity
x <- data.frame(z$WarmingTreatment,z$DivShann)
anova(x)
anovaGraph(x)

# Effect of warming treatment on richness
y <- data.frame(z$WarmingTreatment,z$Richness)
anova(y)
anovaGraph(y)

# Effect of warming treatment on evenness
a <- data.frame(z$WarmingTreatment,z$Evenness)
anova(a)
anovaGraph(a)

```


## 3. Make a random data set that mimics your existing data (test mean and variance) and use the functions

```{r}

# Read in original data
z <- read.table("Data.csv", header = TRUE, sep=",", stringsAsFactors=FALSE)

# Create random response variable data
library(fGarch)

par(mfrow = c(2,1))
hist(rsnorm(120, mean = 2.644833, sd = 0.7477299, xi = -5)) # PLAY AROUND MORE WITH PARAMETERS UNTIL THIS MATCHES SECOND GRAPH
hist(z$DivShann)

mean(z$DivShann)
sd(z$DivShann)





# Test the effect of warming treatment on the response using the random data
randomData <- data.frame(z$WarmingTreatment,y)
anova(randomData)
anovaGraph(randomData)

```


## 4. Manipulate your random data set to mimic a hypothesis you created (i.e. I expect a positive relationship between the number of cats someone owns and that person’s IQ) and use the functions

