---
title: 'Assignment 4'
author: "Melanie Kazenel"
date: "February 8, 2017"
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

#(1) Creating a .csv file in Excel

###Read in the data

```{r}
z <- read.table("Data.csv", header = TRUE, sep=",", stringsAsFactors=FALSE)

```

###Check the data for errors

####str() 
```{r}
str(z)
```

####summary()
```{r}
summary(z)
```

####table()
```{r}
table(z$Sample)
table(z$Plot)
table(z$WarmingTreatment)
table(z$HostSpecies)
table(z$TissueType)
table(z$CollectionDate)
```

####head()
```{r}
head(z)
```

####tail()
```{r}
tail(z)
```


#(2) Working with Regular Expressions

##Question 1

I took this:

```
First String    Second      1.22      3.4
Second          More Text   1.555555  2.2220
Third           x           3         124
```

And transformed it to this:

```
First String,Second,1.22,3.4
Second,More Text,1.555555,2.2220
Third,x,3,124
```

By searching for this:
```
\s{2,}
```

And replacing it with this:
```
,
````

##Question 2

I took this:

```
Ballif, Bryan, University of Vermont
Ellison, Aaron, Harvard Forest
Record, Sydne, Bryn Mawr
```

And transformed it to this:

```
Bryan Ballif (University of Vermont)
Aaron Ellison (Harvard Forest)
Sydne Record (Bryn Mawr)
```

By searching for this:
```
(\w+), (\w+), (.+)
```

And replacing it with this:
```
\2 \1 (\3)
```

##Question 3

###Part 1

I took this:

```
0001 Georgia Horseshoe.mp3 0002 Billy In The Lowground.mp3 0003 Cherokee Shuffle.mp3 0004 Walking Cane.mp3
```

And transformed it to this:

```
0001 Georgia Horseshoe.mp3 
0002 Billy In The Lowground.mp3 
0003 Cherokee Shuffle.mp3 
0004 Walking Cane.mp3
```

By searching for this:
```
mp3 
```

And replacing it with this:
```
mp3 \n
```

###Part 2

I took this:

```
0001 Georgia Horseshoe.mp3 
0002 Billy In The Lowground.mp3 
0003 Cherokee Shuffle.mp3 
0004 Walking Cane.mp3
```

And transformed it to this:

```
Georgia Horseshoe_0001.mp3 
Billy In The Lowground_0002.mp3 
Cherokee Shuffle_0003.mp3 
Walking Cane_0004.mp3
```

By searching for this:
```
(\d+) (.+)[.mp3]{4}
```

And replacing it with this:
```
\2_\1.mp3
```

##Question 4

###Part 1

I took this:

```
Camponotus,pennsylvanicus,10.2,44
Camponotus,herculeanus,10.5,3
Myrmica,punctiventris,12.2,4
Lasius,neoniger,3.3,55
```

And transformed it to this:

```
C_pennsylvanicus,44
C_herculeanus,3
M_punctiventris,4
L_neoniger,55
```

By searching for this:
```
(\w)(\w+),(\w+),(.+),(.+)
```

And replacing it with this:
```
\1_\3,\5
```

###Part 2

I took this:

```
Camponotus,pennsylvanicus,10.2,44
Camponotus,herculeanus,10.5,3
Myrmica,punctiventris,12.2,4
Lasius,neoniger,3.3,55
```

And transformed it to this:

```
C_penn,44
C_herc,3
M_punc,4
L_neon,55
```

By searching for this:
```
(\w)(\w+),(\w{4})(\w+),(.+),(.+)
```

And replacing it with this:
```
\1_\3,\6
```