---
output: 
  beamer_presentation: 
    colortheme: dolphin
    incremental: yes
---
##
\bigskip
\huge Altitudinal gradients do not predict plant-symbiont responses to experimental warming     
\bigskip
\bigskip
\LARGE Melanie Kazenel     
\LARGE 3 May 2017 

 \begin{picture}(320,250)
 \put(140,150){\includegraphics[height=4.5cm]{/Users/mkazenel/Desktop/Bio381/Images/Mountain1.jpg}}
 \put(00,50){\begin{minipage}[t]{1\linewidth}
 \end{minipage}}
 \end{picture}
 
## How will climate change alter species interactions?
 \begin{picture}(320,250)
 \put(40,50){\includegraphics[height=6cm]{/Users/mkazenel/Desktop/Bio381/Images/Mountain2.jpg}}
 \put(25,20){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 
 ## Altitudinal Gradients and Experimental Warming
  \begin{picture}(320,250)
 \put(-5,40){\includegraphics[height=6cm]{/Users/mkazenel/Desktop/Bio381/Images/AltitudeWarming.jpg}}
 \put(25,20){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 
## Questions

- How do fungal symbionts change with altitude and experimental warming?     
- Do these changes correspond?     
 

## Study System and Methods
 
 \begin{picture}(320,250)
 \put(15,30){\includegraphics[height=7cm]{/Users/mkazenel/Desktop/Bio381/Images/PlantSymbiont5.png}}
 \put(25,20){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}

## Statistics

Linear Mixed Effects Models

```{r,eval=FALSE}

library(nlme)

# Altitudinal Gradient Code
alt <- lme(Colonization~Elevation*HostSpecies
              *Year*JulianDay, data = altData, 
              random = ~1|Gradient, method="ML")
anova(alt)

# Experimental Warming Code
warm <- lme(Colonization~WarmingTreatment*HostSpecies
               *Month, data = warmingData, 
               random = ~1|Plot, method="ML")
anova(warm)

```


## Statistics

PERMANOVA

```{r,eval=FALSE}

library(vegan)

adonis(formula = compMatrix~Elevation*HostSpecies
       *Year*JulianDay*Gradient, data = envData, 
       permutations = 999)

```

NMDS

```{r,eval=FALSE}

NMDS <- metaMDS(compMatrix, k=2, try=100, 
                autotransform=TRUE)

```


## Results
- Responses differed among plant species and fungal functional groups    
- Little correspondence of altitude and warming trends    

## Colonization of _Achnatherum lettermanii_ by arbuscular mycorrhizal fungi

```{r, echo=FALSE,results='hide',fig.keep='all',message=FALSE}
library(Rmisc)
library(ggplot2)

#read in data
elevdata <- read.csv("~/Documents/**UNM/Elevation:Warming Work/Stats/ElevationColData.csv")

#change variables to factors
elevdata$Year <- as.factor(elevdata$Year)

#ACLE colonization by AMF 
#create dataset 
dataAMF <- subset(elevdata,Year=="2012" | Year=="2014")
dataAMF <- dataAMF[!is.na(dataAMF$Avg_AMFhyphae),] 
ALdata <- dataAMF[ which(dataAMF$Spp_code=='AL'), ]

p1 <- ggplot(ALdata, aes(x = Elevation, y = Avg_AMFhyphae))
p1 <- p1 + geom_point(colour = "#9e9ac8", size = 5, shape = 16)
p1 <- p1 + xlab("\n Elevation (m)")
#p1 <- p1 + ylab("% Colonization \n")
p1 <- p1 + ylab(NULL)
p1 <- p1 + theme_bw() 
p1 <- p1 + theme(text = element_text(size=25))
p1 <- p1+ theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
p1 <- p1 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
#p1 <- p1 + theme(panel.grid.minor = element_blank())
p1 <- p1 + stat_smooth(method = "lm", formula = y ~ x, size = 0.65, colour = "black", se = FALSE)
#p1 <- p1 + theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())
p1 <- p1 + theme(axis.text.y = element_text(size=20, colour = "black"))
p1 <- p1 + theme(axis.text.x = element_text(size=20, colour = "black"))
p1 <- p1 + xlim(2700, 3500)
p1 <- p1 + ylim(0,100)
p1 <- p1 + theme(axis.text.y = element_blank(),axis.ticks.y = element_blank())
#p1 <- p1 + annotate("text", x = 2715, y = 95, label = "A", size = 10) 
#print(p1)


#AMF colonization, warming
warmingdata <- read.csv("~/Documents/**UNM/Elevation:Warming Work/Stats/warmingdata.csv")
warmingdata$pair <- as.factor(warmingdata$pair)
ALwarming <- subset(warmingdata, hostspecies == "AL")

summary <- summarySE(ALwarming, measurevar="AMFhyphae", groupvars=c("warmingtreatment"))
summary

summary$warmingtreatment <- factor(summary$warmingtreatment, levels = rev(levels(summary$warmingtreatment)))
levels(summary$warmingtreatment)

p2 <- ggplot(summary, aes(warmingtreatment, AMFhyphae))
p2 <- p2 + geom_point(position=position_dodge(width=0.3), stat="identity", size = 5, shape = 16, colour=c("#ca0020","#0571b0")) 
p2 <- p2 + geom_errorbar(aes(ymin=AMFhyphae-se, ymax=AMFhyphae+se),
                       width=.1,position=position_dodge(.3))
p2 <- p2 + xlab("\n  Treatment")
p2 <- p2 + ylab("% Colonization \n")
p2 <- p2 + theme_bw() 
p2 <- p2+ theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
p2 <- p2 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
p2 <- p2+ theme(text = element_text(size=25))
p2 <- p2 + theme(axis.text.x = element_text(size=20, colour = "black"))
p2 <- p2 + theme(axis.text.y = element_text(size=20, colour = "black"))
p2 <- p2 + theme(legend.position="none")
#p2 <- p2 + theme(legend.title=element_blank())
p2 <- p2 + scale_x_discrete(labels = c("Warmed", "Control"))
p2 <- p2 + ylim(0,106)
p2 <- p2 + scale_shape_manual(values=c(16,1))
p2 <- p2+ theme(legend.key = element_blank())
#p2 <- p2 + theme(legend.position = c(.84, .105),legend.background = element_rect(color = "black",size = 0.2))
p2 <- p2 + annotate("text", x = 1.5, y = 80, label = "italic(P)==0.0385", parse = TRUE, size = 6) 
#print(p2)

library(gridExtra)
grid.arrange(p2,p1, ncol=2)



# multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
#   library(grid)
# 
#   # Make a list from the ... arguments and plotlist
#   plots <- c(list(...), plotlist)
# 
#   numPlots = length(plots)
# 
#   # If layout is NULL, then use 'cols' to determine layout
#   if (is.null(layout)) {
#     # Make the panel
#     # ncol: Number of columns of plots
#     # nrow: Number of rows needed, calculated from # of cols
#     layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
#                     ncol = cols, nrow = ceiling(numPlots/cols))
#   }
# 
#  if (numPlots==1) {
#     print(plots[[1]])
# 
#   } else {
#     # Set up the page
#     grid.newpage()
#     pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
# 
#     # Make each plot, in the correct location
#     for (i in 1:numPlots) {
#       # Get the i,j matrix positions of the regions that contain this subplot
#       matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
# 
#       print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
#                                       layout.pos.col = matchidx$col))
#     }
#   }
# }
# 
# multiplot(p2,p1, cols = 2)
```



## Colonization of _Achnatherum lettermanii_ by horizontally transmitted leaf fungi


```{r, echo=FALSE,results='hide',fig.keep='all',message=FALSE}
library(Rmisc)
library(ggplot2)

#read in data
elevdata <- read.csv("~/Documents/**UNM/Elevation:Warming Work/Stats/ElevationColData.csv")

#change variables to factors
elevdata$Year <- as.factor(elevdata$Year)

dataLFE <- subset(elevdata,Year=="2012" | Year=="2014")
dataLFE <- dataLFE[!is.na(dataLFE$AvgOF),] 
ALdataLFE <- dataLFE[ which(dataLFE$Spp_code=='AL'), ]
ALdataLFE$OF <- (ALdataLFE$AvgOF*ALdataLFE$TotalFOV)

p3 <- ggplot(ALdataLFE, aes(x = Elevation, y = OF))
p3 <- p3 + geom_point(colour = "#9e9ac8", size = 5, shape = 16)
p3 <- p3 + xlab("\n Elevation (m)")
p3 <- p3 + ylab(NULL)
#p3 <- p3 + ylab("% Symbiotic Individuals \n")
p3 <- p3 + theme_bw() 
p3 <- p3 + theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
p3 <- p3 + theme(text = element_text(size=25))
p3 <- p3 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
p3 <- p3 + stat_smooth(method = "lm", formula = y ~ x, size = 0.65, colour = "black", se = FALSE)
#p3 <- p3 + theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())
p3 <- p3 + theme(axis.text.y = element_text(size=20, colour = "black"))
p3 <- p3 + theme(axis.text.x = element_text(size=20, colour = "black"))
p3 <- p3 + xlim(2700, 3500)
p3 <- p3 + ylim(0,100)
p3 <- p3 + theme(axis.text.y = element_blank(),axis.ticks.y = element_blank())
p3 <- p3 + annotate("text", x = 2790, y = 9, label = "italic(P)==0.0094", parse = TRUE, size = 6) 
p3 <- p3 + annotate("text", x = 2790, y = 2, label = "italic(r^2)==0.69", parse = TRUE, size = 6) 

#print(p3)


#LFE colonization, warming
warmingdata <- read.csv("~/Documents/**UNM/Elevation:Warming Work/Stats/warmingdata.csv")
warmingdata$pair <- as.factor(warmingdata$pair)
ALwarming <- subset(warmingdata, hostspecies == "AL")

summary <- summarySE(ALwarming, measurevar="meanother", groupvars=c("warmingtreatment"))
summary

summary$warmingtreatment <- factor(summary$warmingtreatment, levels = rev(levels(summary$warmingtreatment)))
levels(summary$warmingtreatment)

p4 <- ggplot(summary, aes(warmingtreatment, meanother))
p4 <- p4 + geom_point(position=position_dodge(width=0.3), stat="identity", size = 5, shape = 16, colour=c("#ca0020","#0571b0")) 
p4 <- p4 + geom_errorbar(aes(ymin=meanother-se, ymax=meanother+se),
                       width=.1,position=position_dodge(.3))
p4 <- p4 + xlab("\n  Treatment")
p4 <- p4 + ylab("% Colonization \n")
p4 <- p4 + theme_bw() 
p4 <- p4+ theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
p4 <- p4 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
p4 <- p4+ theme(text = element_text(size=25))
p4 <- p4 + theme(axis.text.x = element_text(size=20, colour = "black"))
p4 <- p4 + theme(axis.text.y = element_text(size=20, colour = "black"))
p4 <- p4 + theme(legend.title=element_blank())
p4 <- p4 + scale_x_discrete(labels = c("Warmed", "Control"))
p4 <- p4 + ylim(0,100)
p4 <- p4 + scale_shape_manual(values=c(16,1))
p4 <- p4+ theme(legend.key = element_blank())
#print(p4)

grid.arrange(p4,p3, ncol=2)

# multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
#   library(grid)
# 
#   # Make a list from the ... arguments and plotlist
#   plots <- c(list(...), plotlist)
# 
#   numPlots = length(plots)
# 
#   # If layout is NULL, then use 'cols' to determine layout
#   if (is.null(layout)) {
#     # Make the panel
#     # ncol: Number of columns of plots
#     # nrow: Number of rows needed, calculated from # of cols
#     layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
#                     ncol = cols, nrow = ceiling(numPlots/cols))
#   }
# 
#  if (numPlots==1) {
#     print(plots[[1]])
# 
#   } else {
#     # Set up the page
#     grid.newpage()
#     pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
# 
#     # Make each plot, in the correct location
#     for (i in 1:numPlots) {
#       # Get the i,j matrix positions of the regions that contain this subplot
#       matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
# 
#       print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
#                                       layout.pos.col = matchidx$col))
#     }
#   }
# }
# 
# multiplot(p4,p3, cols = 2)
```


## Colonization of _Festuca thurberi_ by vertically transmitted leaf fungi

```{r, echo=FALSE,results='hide',fig.keep='all',message=FALSE,warning=FALSE}
library(Rmisc)
library(ggplot2)

#read in data
elevdata <- read.csv("~/Documents/**UNM/Elevation:Warming Work/Stats/ElevationColData.csv")

#change variables to factors
elevdata$Year <- as.factor(elevdata$Year)

data2 <- subset(elevdata, Year=="2008" | Year=="2009" | Year=="2011" | Year=="2012" | Year=="2014")
data2 <- data2[!is.na(data2$AvgEndo),]  
FTdata <- data2[ which(data2$Spp_code=='FT'), ]
FTdata$Endo <- (FTdata$AvgEndo*FTdata$TotalFOV)

p3 <- ggplot(FTdata, aes(x = Elevation, y = Endo))
p3 <- p3 + geom_point(colour = "#9e9ac8", size = 5, shape = 16)
p3 <- p3 + xlab("\n Elevation (m)")
p3 <- p3 + ylab(NULL)
#p3 <- p3 + ylab("% Symbiotic Individuals \n")
p3 <- p3 + theme_bw() 
p3 <- p3 + theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
p3 <- p3 + theme(text = element_text(size=25))
p3 <- p3 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
p3 <- p3 + stat_smooth(method = "lm", formula = y ~ x, size = 0.65, colour = "black", se = FALSE)
#p3 <- p3 + theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())
p3 <- p3 + theme(axis.text.y = element_text(size=20, colour = "black"))
p3 <- p3 + theme(axis.text.x = element_text(size=20, colour = "black"))
p3 <- p3 + xlim(2700, 3500)
p3 <- p3 + ylim(0,100)
p3 <- p3 + theme(axis.text.y = element_blank(),axis.ticks.y = element_blank())
p3 <- p3 + annotate("text", x = 3400, y = 90, label = "italic(P)==0.0023", parse = TRUE, size = 6) 
p3 <- p3 + annotate("text", x = 3400, y = 83, label = "italic(r^2)==0.33", parse = TRUE, size = 6) 

#print(p3)



#Epicolonization, warming
warmingdata <- read.csv("~/Documents/**UNM/Elevation:Warming Work/Stats/warmingdata.csv")
warmingdata$pair <- as.factor(warmingdata$pair)
FTwarming <- subset(warmingdata, hostspecies == "FT")

summary <- summarySE(FTwarming, measurevar="meanepi", groupvars=c("warmingtreatment"))
summary

summary$warmingtreatment <- factor(summary$warmingtreatment, levels = rev(levels(summary$warmingtreatment)))
levels(summary$warmingtreatment)

p4 <- ggplot(summary, aes(warmingtreatment, meanepi))
p4 <- p4 + geom_point(position=position_dodge(width=0.3), stat="identity", size = 5, shape = 16, colour=c("#ca0020","#0571b0")) 
p4 <- p4 + geom_errorbar(aes(ymin=meanepi-se, ymax=meanepi+se),
                       width=.1,position=position_dodge(.3))
p4 <- p4 + xlab("\n  Treatment")
p4 <- p4 + ylab("% Colonization \n")
p4 <- p4 + theme_bw() 
p4 <- p4+ theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
p4 <- p4 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
p4 <- p4+ theme(text = element_text(size=25))
p4 <- p4 + theme(axis.text.x = element_text(size=20, colour = "black"))
p4 <- p4 + theme(axis.text.y = element_text(size=20, colour = "black"))
p4 <- p4 + theme(legend.title=element_blank())
p4 <- p4 + scale_x_discrete(labels = c("Warmed", "Control"))
p4 <- p4 + ylim(0,100)
p4 <- p4 + scale_shape_manual(values=c(16,1))
p4 <- p4+ theme(legend.key = element_blank())
#p4 <- p4 + theme(legend.position = c(.84, .105),legend.background = element_rect(color = "black",size = 0.2))
p4 <- p4 + annotate("text", x = 1.5, y = 50, label = "italic(P)==0.0673", parse = TRUE, size = 6) 

#print(p4)

grid.arrange(p4,p3, ncol=2)

# multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
#   library(grid)
# 
#   # Make a list from the ... arguments and plotlist
#   plots <- c(list(...), plotlist)
# 
#   numPlots = length(plots)
# 
#   # If layout is NULL, then use 'cols' to determine layout
#   if (is.null(layout)) {
#     # Make the panel
#     # ncol: Number of columns of plots
#     # nrow: Number of rows needed, calculated from # of cols
#     layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
#                     ncol = cols, nrow = ceiling(numPlots/cols))
#   }
# 
#  if (numPlots==1) {
#     print(plots[[1]])
# 
#   } else {
#     # Set up the page
#     grid.newpage()
#     pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
# 
#     # Make each plot, in the correct location
#     for (i in 1:numPlots) {
#       # Get the i,j matrix positions of the regions that contain this subplot
#       matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
# 
#       print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
#                                       layout.pos.col = matchidx$col))
#     }
#   }
# }
# 
# multiplot(p4,p3, cols = 2)
```


## Possible Explanations for Lack of Correspondence in Altitude and Warming Trends

- Factors other than temperature drive altitudinal patterns    
- Gradients captured a broader range of temperatures    
- In different parts of their ranges, organisms can respond differently to warming    
- Experimental warming represents an immediate step change in temperature    



## Fungal community composition did not respond to altitude or warming

```{r, echo=FALSE,results='hide',fig.keep='all',message=FALSE,warning=FALSE}

# ROOTS
# library(vegan)
# z <- read.csv("WarmingRoots.csv")
# z$Sample <- NULL
# NMDS <- metaMDS(z, k=2, trymax=100, autotransform=TRUE)
# 
# library(ggplot2)
# scores <- as.data.frame(scores(NMDS))
# env <- read.csv("WarmingEnv.csv")
# envRoots <- subset(env,TissueType=="Rts")
# scores$warming <- envRoots$WarmingTreatment
# scores$date <- envRoots$CollectionDate
# print(scores)

#write.csv(scores,"scores.csv")

scores <- read.csv("scores.csv")

scores$warming <- factor(scores$warming, levels = rev(levels(scores$warming)))
levels(scores$warming)

#####p4 <- ggplot(summary, aes(warmingtreatment, meanepi))
####p4 <- p4 + geom_point(position=position_dodge(width=0.3), stat="identity", size = 5, ####shape = 16, colour=c("#ca0020","#0571b0")) 


p <- ggplot(scores, aes(NMDS1, NMDS2, colour=warming))
p <- p + geom_point(size = 5, shape = 16)
p <- p + theme_bw() 
p <- p + theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
p <- p + theme(text = element_text(size=25))
p <- p + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
p <- p + theme(axis.text.y = element_text(size=20, colour = "black"))
p <- p + theme(axis.text.x = element_text(size=20, colour = "black"))
p <- p + theme(legend.title=element_blank())
#p <- p + theme(legend.position = "none")
p <- p + ggtitle("Root Fungi")
p <- p + theme(legend.position="none")
p <- p + xlim(-1.5,1.5)
p <- p + ylim(-1.5,1.5)
#p4 <- p4 + theme(legend.key = element_blank())
#p <- p + theme(legend.background = element_rect(color = "black",size = 0.2)) #legend.position = c(.9, .785),
#print(p)


#LEAVES 

# z2 <- read.csv("WarmingLeaves.csv")
# z2$Sample <- NULL
# NMDSleaves <- metaMDS(z2, k=2, trymax=100, autotransform=TRUE)
# 
# scoresLeaves <- as.data.frame(scores(NMDSleaves))
# env <- read.csv("WarmingEnv.csv")
# envLeaves <- subset(env,TissueType=="Lvs")
# scoresLeaves$warming <- envLeaves$WarmingTreatment
# scoresLeaves$date <- envLeaves$CollectionDate
# print(scoresLeaves)

# write.csv(scoresLeaves,"scoresLeaves.csv")

scoresLeaves <- read.csv("scoresLeaves.csv")

scoresLeaves$warming <- factor(scoresLeaves$warming, levels = rev(levels(scoresLeaves$warming)))
levels(scoresLeaves$warming)

p2 <- ggplot(scoresLeaves, aes(x = NMDS1, y = NMDS2, colour=warming))
p2 <- p2 + geom_point(size = 5, shape = 16)
p2 <- p2 + theme_bw() 
p2 <- p2 + theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
p2 <- p2 + theme(text = element_text(size=25))
p2 <- p2 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
p2 <- p2 + theme(axis.text.y = element_text(size=20, colour = "black"))
p2 <- p2 + theme(axis.text.x = element_text(size=20, colour = "black"))
p2 <- p2 + theme(legend.title=element_blank())
p2 <- p2 + ggtitle("Leaf Fungi")
p2 <- p2 + xlim(-1.5,1.5)
p2 <- p2 + ylim(-1.5,1.5)
#p4 <- p4 + theme(legend.key = element_blank())
#p <- p + theme(legend.background = element_rect(color = "black",size = 0.2)) #legend.position = c(.9, .785),
#print(p2)

library(gridExtra)
# get_legend<-function(myggplot){
#   tmp <- ggplot_gtable(ggplot_build(myggplot))
#   leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
#   legend <- tmp$grobs[[leg]]
#   return(legend)
# }
# legend <- get_legend(p)
#p <- p + theme(legend.position="none")
#p2 <- p2 + theme(legend.position="none")

grid.arrange(p, p2, ncol=2, widths=c(2,2.75), heights=c(2,2))

#multiplot(p,p2,cols=2)

```




## Fungal community composition varied across the growing season

```{r, echo=FALSE,results='hide',fig.keep='all',message=FALSE,warning=FALSE}

# ROOTS by date

scores <- read.csv("scores.csv")

p1 <- ggplot(scores, aes(x = NMDS1, y = NMDS2, colour=date))
p1 <- p1 + geom_point(size = 5, shape = 16)
p1 <- p1 + scale_color_brewer(palette="Set2")
#p <- p + xlab("\n Elevation (m)")
#p <- p + ylab("% Symbiotic Individuals \n")
p1 <- p1 + theme_bw() 
p1 <- p1 + theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
p1 <- p1 + theme(text = element_text(size=25))
p1 <- p1 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
p1 <- p1 + theme(axis.text.y = element_text(size=20, colour = "black"))
p1 <- p1 + theme(axis.text.x = element_text(size=20, colour = "black"))
p1 <- p1 + theme(legend.title=element_blank())
p1 <- p1 + ggtitle("Root Fungi")
p1 <- p1 + theme(legend.position="none")
p1 <- p1 + xlim(-1.5,1.5)
p1 <- p1 + ylim(-1.5,1.5)
p1 <- p1 + annotate("text", x = 1, y = -1.3, label = "italic(P)<0.001", parse = TRUE, size = 6) 
#p4 <- p4 + theme(legend.key = element_blank())
#p <- p + theme(legend.background = element_rect(color = "black",size = 0.2)) #legend.position = c(.9, .785),
#print(p1)


# Leaves by date

scoresLeaves <- read.csv("scoresLeaves.csv")

p <- ggplot(scoresLeaves, aes(x = NMDS1, y = NMDS2, colour=date))
p <- p + geom_point(size = 5, shape = 16)
p <- p + scale_color_brewer(palette="Set2")
#p <- p + xlab("\n Elevation (m)")
#p <- p + ylab("% Symbiotic Individuals \n")
p <- p + theme_bw() 
p <- p + theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
p <- p + theme(text = element_text(size=25))
p <- p + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
p <- p + theme(axis.text.y = element_text(size=20, colour = "black"))
p <- p + theme(axis.text.x = element_text(size=20, colour = "black"))
p <- p + theme(legend.title=element_blank())
p <- p + ggtitle("Leaf Fungi")
p <- p + xlim(-1.5,1.5)
p <- p + ylim(-1.5,1.5)
p <- p + annotate("text", x = 1, y = -1.3, label = "italic(P)<0.001", parse = TRUE, size = 6) 
#p4 <- p4 + theme(legend.key = element_blank())
#p <- p + theme(legend.background = element_rect(color = "black",size = 0.2)) #legend.position = c(.9, .785),
#print(p)

grid.arrange(p1, p, ncol=2, widths=c(2,2.9), heights=c(2,2))

#multiplot(p1,p,cols=2)

```


## Conclusions

- Responses of symbioses to climate change will differ among plant species and fungal functional groups    
- Little correspondence of altitude and warming patterns    
- Climate change may affect fungal colonization more strongly than composition    
- Potential for phenological mismatches under climate change    

## Acknowledgments
>- Jennifer Rudgers    
>- D. Lee Taylor, Cristina Takacs-Vesbach    
>- Stephanie Kivlin     
>- Joshua Lynn    
>- Rocky Mountain Biological Laboratory and University of New Mexico undergraduates    
\bigskip
>- National Science Foundation (DEB-1354972)    
>- Rocky Mountain Biological Laboratory (NSF DBI-0753774 and OIA-0963529)    
>- University of New Mexico Department of Biology    


## Questions?
\bigskip
  \begin{picture}(320,250)
 \put(50,30){\includegraphics[height=7.5cm]{/Users/mkazenel/Desktop/Bio381/Images/FinalPhoto.png}}
 \put(20,40){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture} 
 