---
title: "'plyr' tutorial"
author: "Melanie Kazenel"
date: "19 April 2017"
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

Required file:
This tutorial uses publically available climate data collected by two different weather stations in the Rocky Mountains of Colorado, USA. The dataset is available [here](ClimateData.csv).

### Install the required package and read in the data

```{r}
#install.packages("plyr")
library(plyr)

climate <- read.csv(file="ClimateData.csv")
climate$Month <- as.factor(climate$Month)
climate$Year <- as.factor(climate$Year)
str(climate)
```

### Using 'ddply': a simple example
```{r}
# Use 'ddply' to calculate the mean air temperature for each month
ddply(climate, # input data frame
      "Month", # variable to subset by
      function(x){ # function to run on each subset
        mean(x$AvgAirTemp)
        }
      )

# Alter the function slightly so that the output is easier to work with
monthlyData <- ddply(climate,
                     "Month",
                     function(x){ 
                       MeanAirTemp <- mean(x$AvgAirTemp)
                       data.frame(MeanAirTemp=MeanAirTemp)
                       }
                     )
print(monthlyData)
```

### Using 'ddply' to subset data based on multiple factors and perform a calculation
```{r}
# Use 'ddply' to calculate the mean air temperature for each month-year combination
monthYearData <- ddply(climate,
                     c("Month","Year"),
                     function(x){ 
                       MeanAirTemp <- mean(x$AvgAirTemp)
                       data.frame(MeanAirTemp=MeanAirTemp)
                       }
                     )
print(monthYearData)
```


### Using 'ddply' to calculate multiple summary statistics
```{r}
# Use 'ddply' to calculate, for each month, means and standard deviations for daily air
# temperature and precipitation
monthlyData <- ddply(climate,
                     "Month",
                     function(x){ 
                       
                       meanAirTemp <- mean(x$AvgAirTemp)
                       sdAirTemp <- sd(x$AvgAirTemp)
                       meanPrecip <- mean(x$Precip)
                       sdPrecip <- sd(x$Precip)
                       
                       data.frame(meanAirTemp=meanAirTemp,sdAirTemp=sdAirTemp,
                                  meanPrecip=meanPrecip,sdPrecip=sdPrecip)
                       }
                     )
print(monthlyData)
```

### Using 'summarise' within 'ddply'
```{r}
# Calculate the mean air temperature for each month
monthlyData <- ddply(climate, # input data frame 
           "Month", # variable to subset by
           summarise, # "helper function" to run
           MeanAirTemp = mean(AvgAirTemp)) # function to apply to each subset
print(monthlyData)

# Calculate the mean air temperature for each month-year combination
monthYearData <- ddply(climate, 
           c("Month","Year"), 
           summarise, 
           MeanAirTemp = mean(AvgAirTemp))
print(monthYearData)

# Calculate, for each month, the mean and standard deviation for air temperature 
monthlyData <- ddply(climate, 
           "Month", 
           summarise, 
           meanAirTemp=mean(AvgAirTemp), 
           sdAirTemp=sd(AvgAirTemp))
print(monthlyData)
```

### Using 'transform' and 'mutate' within 'ddply'
```{r}
# Use 'transform' within 'ddply' to split your data into subsets, perform a calculation on
# each subset, and add the results to a copy of your input data frame as a new column
x <- ddply(climate, 
           "Month", 
           transform, 
           MonthlyMeanTemp = mean(AvgAirTemp))
head(x)

# 'Mutate' works similarly to 'transform', but allows you to do calculations within 'ddply'
# using columns you just created
x <- ddply(climate, 
           "Month", 
           mutate, 
           AvgMaxTemp = mean(MaxAirTemp),
           AvgMinTemp = mean(MinAirTemp),
           MonthlyMeanTempRange = AvgMaxTemp - AvgMinTemp)
head(x)
```

### Using plyr for plotting
```{r}
# Create boxplots of the mean daily air temperatures for each weather station
par(mfrow = c(1,2))
d_ply(climate, 
      "StationName", 
      summarise, 
      boxplot(AvgAirTemp, 
              xlab=unique(StationName), 
              ylab="Mean Daily Air Temperature (degrees C)"))
```

### Using other types of functions within 'ddply'
```{r}
# Use a linear model to examine how precipitation changes with air temperature within 
# each month
precipTemp <- ddply(climate, 
                     "Month", 
                     function(x) {
                       model <- lm(Precip ~ AvgAirTemp, data=x)
                       setNames(coef(model), c("Intercept", "Slope"))
                       }
                     )
print(precipTemp)

```

### Creating graphs using 'ddply' output

```{r}
# For each month, calculate and graph the mean and standard deviation for air temperature 

monthlyData <- ddply(climate, 
           "Month", 
           summarise, 
           meanAirTemp=mean(AvgAirTemp), 
           sdAirTemp=sd(AvgAirTemp))
print(monthlyData)

library(ggplot2)

p <- ggplot(monthlyData, aes(x=Month,y=meanAirTemp,colour=Month))
p <- p + geom_point(position=position_dodge(width=0.3), stat="identity", size = 3) 
p <- p + geom_errorbar(aes(ymin=meanAirTemp-sdAirTemp, ymax=meanAirTemp+sdAirTemp),
                         width=.1,position=position_dodge(.3))
print(p)

```

