---
title: "Working with Spatial Data and using tmap()"
author: "Samantha A. Alger"
date: "4/26/2017"
output: html_document
---
Modified from:

* https://cran.r-project.org/doc/contrib/intro-spatial-rl.pdf

* https://github.com/mtennekes/tmap/tree/master/demo/BubbleMap)


###INSTALL PACKAGES and SET WORKING DIRECTORY

```{r results='hide', message=FALSE, warning=FALSE}


rm(list=ls())

x <- c("rgdal","tmap")
#install.packages(x)
lapply(x, library, character.only = TRUE)

setwd("~/CompBio/MapTutorial/") # setting working directory

```

###STRUCTURE OF SPATIAL DATA IN R

Spatial objects such as lnd are made up of different 'slots', key slots being '@data' (non geographic attribute data) and @polygons (or @lines for line data) (polygons that make up the physicial boundaries). Specific slots are accessed using the '@' symbol. Using the '$' symbol to see attribute in the @ data slot. 

```{r results='hide', message=FALSE, warning=FALSE}
#Read in the data:

lnd <- readOGR("data", "london_sport") # Read in the data

head(lnd@data) # use the @ symbol to refer to the data slot of the lnd object

mean(lnd$Partic_Per) # use the $ symbol to refer to an attribute in the 'data' slot

sapply(lnd@data, class) # To check the classes of all the variabes in a spatial dataset, you can use this command

lnd$Pop_2001 <- as.numeric(as.character(lnd$Pop_2001)) # Unexpectedly, Pop_2001 is a factor, coerce the variable into the correct, numeric format...

# type "lnd@" and then tab to see which slots are in lnd...

#check number of rows and columns...
nrow(lnd)
ncol(lnd)
```

###BASIC PLOTTING

```{r}

plot(lnd)
```

###DATA QUERYING AND PLOTTING
```{r}
# select rows of lnd@data where sports participation is less than 15
lnd@data[lnd$Partic_Per < 15, ] # This is interrogating the attribute data slot only of the lnd object..

#Using the same logic to subset and plot spatial objects (polygon slot)

# Select zones where sports participation is between 20 and 25%
sel <- lnd$Partic_Per > 20 & lnd$Partic_Per < 25
head(sel) # test output of previous selection
plot(lnd[sel, ]) # plotting all TRUE Cases

# To see these areas in context with other areas of the maps... use add= TRUE..

plot(lnd, col = "lightgrey") # plot the london_sport object
sel <- lnd$Partic_Per > 25
plot(lnd[ sel, ], col = "turquoise", add = TRUE) # add selected zones to map

#Areas of high sports participation are now highlighted in blue!
```

###USING tmap()
Thematic maps are geographical maps in which spatial data distributions are visualized. This package offers a flexible, layer-based, and easy to use approach to create thematic maps, such as choropleths and bubble maps.

```{r}
library(tmap) # load tmap package

tmap_mode="plot"

# Check out it's capabilities!
#vignette("tmap-nutshell")
```

###Quick Thematic Maps (qtm):
```{r}
# load spatial data included in the tmap package
data("World", "metro")

# Try typing head(World$...) or head(metro@...) to see the type of data...
head(World@data)
head(metro@data)

#  quick plot example: 
qtm(World, fill = "income_grp", text = "iso_a3", text.size = "AREA") # use "World$" to see the two attributes: income_grp and iso_a3, text.size= area: text is sized increasingly with coutry area size.

```

###MORE COMPLEX MAPPING

For More Complex Maps use the main plotting method for readability
```{r}
# calculate annual growth rate
metro$growth <- (metro$pop2020 - metro$pop2010) / (metro$pop2010 * 10) * 100

# plot
tm_shape(World) +
	tm_polygons("income_grp", palette = "-Blues", 
		title = "Income class", contrast = 0.7, border.col = "grey30", id = "name") +

	tm_text("iso_a3", size = "AREA", col = "grey30", root=3) +

  tm_shape(metro) +
  tm_bubbles("pop2010", col = "growth", border.col = "black",
		border.alpha = 0.5,
		breaks = c(-Inf, 0, 2, 4, 6, Inf) ,
		palette = "-RdYlGn",
		title.size = "Metro population (2010)", 
		title.col = "Annual growth rate (%)",
		id = "name",
		popup.vars=c("pop2010", "pop2020", "growth")) + 
tm_format_World() + 
tm_style_gray()

```

###INTERACTIVE VIEW MODE
Interactive viewing with tmap
```{r}
tmap_mode("view") # sets the tmap mode to interactive viewing
last_map() # returns the last map to be modified or created
```
